---
title: "NU tet pos omicron analysis - hetero and homo responses - Sheffield"
author: "Rebecca Payne"
date: "14/11/2022"
output:
  pdf_document: default
  html_document: default
---
Analysis of antigen specific responses from nine batches. Samples are either stained for EBV, CMV, or Spike/non Spike pentamer responses. Pre-gated fcs files on live, CD8+ tetramer positive cells. Cut-off greater than 0.01% response applied

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(readxl) 
library(CATALYST)
library(cowplot)
library(flowCore)
library(scater)
library(SingleCellExperiment)
library(openxlsx)
library(diffcyt)
library(ggpubr)
```



Load  FCS files (processed), metadata, panel and create flowset

```{r}
md <- "metadata_aurora_sheffield.xlsx" 
md <- read_excel(md)       
head(data.frame(md)) 
```


```{r message=FALSE}
fs <- read.flowSet(md$file_name, transformation = F, truncate_max_range = F)
```

```{r}
panel <- "Panel_aurora_sheffield.xlsx" 
panel <- read_excel(panel)                                         
head(data.frame(panel),30) 
```


```{r}
all(panel$fcs_colname %in% colnames(fs))
```

```{r}
setdiff(fs@frames$CD8_Tet2_RPP_APC_3108_V7.fcs@parameters@data$desc,panel$antigen)
```


```{r}
md$infection <- factor(md$infection, levels = c("pre_omicron", "post_omicron"))

md$batch <- factor(md$batch, levels = c("3", "4", "5", "6","7", "8", "9"))

md$antigen <- factor(md$antigen, levels = c("spike", "nonspike", "EBV", "CMV" ))

md$hla <- factor(md$hla, levels = c("B5701", "A0201", "A0301", "B0702", "A0101"))

md$serostatus <- factor(md$serostatus, levels = c("naive", "conv"))





md$condition <- factor(md$condition, levels = c("pre_omicron_spike_naive", "post_omicron_spike_naive", "pre_omicron_spike_previously_infected", "post_omicron_spike_previously_infected", "post_omicron_nonspike_naive", "pre_omicron_nonspike_previously_infected", "post_omicron_nonspike_previously_infected", "pre_omicron_EBV", "post_omicron_EBV", "pre_omicron_CMV", "post_omicron_CMV"))

md$sample_id <- factor(md$sample_id, levels = md$sample_id[order(md$condition)])


```


```{asis, echo = TRUE}
# see ?prepdata for notes on the sce. Aurora flow analysis needs a different cofactor applied. We have chosen 150. Also need to specify FACE = TRUE for flow data.
```

```{r}
sce <- prepData(fs, panel, md, transform = TRUE, cofactor = 150, FACS = TRUE, features = panel$fcs_colname, md_cols = list(file = "file_name", id = "sample_id", factors = c("batch", "condition", "antigen", "hla", "serostatus", "infection")))

```



## Overview analysis of Cytof data

```{r}
n_cells(sce)
```

```{r fig.width= 10}
plotCounts(sce, color_by = "antigen")
```



```{r fig.width= 5}
plotNRS(sce, features = type_markers(sce), color_by = "condition")

```


## Clustering of fcs files with over 0.01% of CD8 (or more than 20 events) tetramer positive events  


### sce_all : Cluster of sce on type AND state markers; using features = NULL
Cluster FLOWSom analysis - max 10 clusters <br/> cluster on "type" markers for memory phenotyping

For reproducibility set random seed
```{r}
set.seed(4321)
sce_all <- cluster(sce, features = NULL, 
               xdim = 10, ydim = 10, maxK = 10, 
               verbose = FALSE, seed = 4321) 
```


```{r fig.width=5}
pbMDS(sce_all, shape_by = "antigen", size_by = TRUE, label_by = NULL)
```

## Dimension reduction - UMAP
```{r}
set.seed(4321)
sce_all <- runDR(sce_all, dr = "UMAP", cells = 500, features = NULL)
```

```{r}
plotDR(sce_all, "UMAP", color_by = "meta10")

```





## 28.11.2022 Group data in sce_all into three groups

## group_a - pre_omicron - only pre_omicron samples 
## group_b - spike only samples
## group_c - nonspike only 

```{r}
sce_all_groupa <- filterSCE(sce_all, infection == "pre_omicron")
```


```{r}
sce_all_groupb <- filterSCE(sce_all, antigen == "spike")
```


```{r}
sce_all_groupc <- filterSCE(sce_all, antigen == "nonspike")
```

## Plots of UMAP based on FLowSOM clustering and DR on sce_all (all markers inc type and state)

```{r}
plotDR(sce_all, "UMAP", color_by = "condition")

```


```{r}
plotDR(sce_all_groupa, "UMAP", color_by = "condition")

```

```{r}
plotDR(sce_all_groupb, "UMAP", color_by = "condition")

```

```{r}
plotDR(sce_all_groupc, "UMAP", color_by = "condition")

```
## Heatmaps


```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_all, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```


```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_all_groupa, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```

```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_all_groupb, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```

```{r fig.width=7, fig.height=5}
plotExprHeatmap(sce_all_groupc, features = NULL, 
    scale = "last", q = 0, bars = FALSE, row_anno = c("condition", "hla"), row_clust = FALSE)
```

# Marker expression 



```{r fig.height=8}
plotPbExprs(sce_all_groupa, features=NULL, facet_by = "antigen", ncol = 3) + stat_compare_means(label.y = -3) + stat_compare_means(method = "t.test", ref.group = ".all.")
```

```{r fig.height=3}
plotPbExprs(sce_all_groupa, features=NULL, facet_by = "antigen", ncol = 3)
```

```{r fig.height=8}
plotPbExprs(sce_all_groupb, features=NULL, facet_by = "antigen", ncol = 3 ) + stat_compare_means(label.y = -3) + stat_compare_means(method = "t.test", ref.group = ".all.")
```


```{r fig.height=3}
plotPbExprs(sce_all_groupb, features=NULL, facet_by = "antigen", ncol = 3 ) 
```

```{r fig.height=8}
plotPbExprs(sce_all_groupc, features=NULL, facet_by = "antigen", ncol = 3) + stat_compare_means(label.y = -3) + stat_compare_means(method = "t.test", ref.group = ".all.", label="p.signif")
```

```{r fig.height=3}
plotPbExprs(sce_all_groupc, features=NULL, facet_by = "antigen", ncol = 3) 
```

```{r}
sessionInfo()
```






